Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-10 csf-2017/paper.tex
he/she to they

The final semantics we discuss is that of $\store{}{}$. As expected, it updates
the memory to the requested value if the capability has write authority for the
requested location. However, the instruction is complicated by the presence of
\emph{local capabilities}, modeled after the ones in
CHERI~\cite{Watson2015Cheri}. Basically, local capabilities are special in that
they can only be kept in registers, i.e. they cannot be stored to memory. This
means that local capabilities can be \emph{temporarily} given to an adversary,
for the duration of an invocation: if we take care to erase the capability from
the register file after control is passed back to us, we can be sure that he/she
will not have been able to store the capability. However, there is one exception
to the rule above: local capabilities can be stored to a part of memory for
which we have a capability with write-local authority (i.e. permission $\rwl$ or
$\rwlx$). This exception is intended to accomodate regions of memory that behave
like the C stack: stashing away the contents of registers on the stack can be
allowed, even if this includes local capabilities. As long as all capabilities
with write-local authority are themselves local and the stack is cleared after
control is passed back by the adversary, we will see that this does not break
the intended behavior of local capabilities.

@@

\textbf{Returning to the adversary} Perhaps the simplest scenario is the case of
returning to the adversary after he/she invoked our code. In this case, we will
have received a return pointer from him, and we just need to jump to it as
usual. The only thing we need to take care of is to properly clear the registers
before we do (erasing any data we do not want to pass as (part of) the return
value) and similarly, erase any part of the stack that we have used (i.e.
overwrite it with zeros). Note that we can use the stack as normal: we may push
values onto it and pop them back off (e.g. for register spilling etc.).

@@

Because the return pointer is local, this is not possible: the adversary can
only store it through write-local capabilities, which (because of our
assumptions above) means: on the stack. However, this makes it crucial that the
stack pointer is also local: if the adversary were able to store the stack
pointer in a global data structure, he/she can store return pointers on the
stack, and play similar mischief as mentioned above. Storing away stack pointers
for use during future invocations would also be dangerous for other reasons:
during such a future invocation, a larger part of the stack may be in use by the
trusted code, which the adversary may still have access to through the old stack
pointer.

@@

It's important that neither of these capabilities are accessible to the
adversary. In other words, the return pointer provided to the adversary must be
a capability that he/she can jump to but not read from, i.e. an enter
capability. To make this work, we construct the activation record as depicted in
Figure~\ref{fig:activ-rec-struct}. The return pointer is an $\entry$ capability
with authority over the activation record containing the previous return and
stack pointer as well as a number of restore instructions. The return pointer's
current address points to those instructions, so that, upon invocation, they are
executed and can load the old stack pointer and program counter back into the
register file. Because the return pointer is given to the adversary as an enter
pointer, he/she cannot get hold of the contents of the data structure, but after
invocation, its permission is updated to $\exec$, so the activation record's
contents become available to the restore instructions.

@@

\textbf{Invoking an adversary callback} If we have a higher-order interface to
the adversary, we might need to invoke an adversary callback. In this case, not
so much changes with respect to the situation where we invoke regular adversary
code. The adversary can provide a callback as a capability for us to jump to,
either an $\entry$-capability if he wants to protect himself from us or just an
$\exec$ capability if he/she is not worried about that. However, there is one
scenario that we need to prevent, namely that he/she constructs the callback
capability to point into the stack, where he/she may have stored local
capabilities that he should not have access to upon invocation of the callback.
Specifically, he might store return pointers and stack pointers from previous
stack frames there, and illegally use them inside the callback.

>>>>

The final semantics we discuss is that of $\store{}{}$. As expected, it updates
the memory to the requested value if the capability has write authority for the
requested location. However, the instruction is complicated by the presence of
\emph{local capabilities}, modeled after the ones in
CHERI~\cite{Watson2015Cheri}. Basically, local capabilities are special in that
they can only be kept in registers, i.e. they cannot be stored to memory. This
means that local capabilities can be \emph{temporarily} given to an adversary,
for the duration of an invocation: if we take care to erase the capability from
the register file after control is passed back to us, we can be sure that they
will not have been able to store the capability. However, there is one exception
to the rule above: local capabilities can be stored to a part of memory for
which we have a capability with write-local authority (i.e. permission $\rwl$ or
$\rwlx$). This exception is intended to accomodate regions of memory that behave
like the C stack: stashing away the contents of registers on the stack can be
allowed, even if this includes local capabilities. As long as all capabilities
with write-local authority are themselves local and the stack is cleared after
control is passed back by the adversary, we will see that this does not break
the intended behavior of local capabilities.

@@

\textbf{Returning to the adversary} Perhaps the simplest scenario is the case of
returning to the adversary after they invoked our code. In this case, we will
have received a return pointer from him, and we just need to jump to it as
usual. The only thing we need to take care of is to properly clear the registers
before we do (erasing any data we do not want to pass as (part of) the return
value) and similarly, erase any part of the stack that we have used (i.e.
overwrite it with zeros). Note that we can use the stack as normal: we may push
values onto it and pop them back off (e.g. for register spilling etc.).

@@

Because the return pointer is local, this is not possible: the adversary can
only store it through write-local capabilities, which (because of our
assumptions above) means: on the stack. However, this makes it crucial that the
stack pointer is also local: if the adversary were able to store the stack
pointer in a global data structure, they can store return pointers on the
stack, and play similar mischief as mentioned above. Storing away stack pointers
for use during future invocations would also be dangerous for other reasons:
during such a future invocation, a larger part of the stack may be in use by the
trusted code, which the adversary may still have access to through the old stack
pointer.

@@

It's important that neither of these capabilities are accessible to the
adversary. In other words, the return pointer provided to the adversary must be
a capability that they can jump to but not read from, i.e. an enter
capability. To make this work, we construct the activation record as depicted in
Figure~\ref{fig:activ-rec-struct}. The return pointer is an $\entry$ capability
with authority over the activation record containing the previous return and
stack pointer as well as a number of restore instructions. The return pointer's
current address points to those instructions, so that, upon invocation, they are
executed and can load the old stack pointer and program counter back into the
register file. Because the return pointer is given to the adversary as an enter
pointer, they cannot get hold of the contents of the data structure, but after
invocation, its permission is updated to $\exec$, so the activation record's
contents become available to the restore instructions.

@@

\textbf{Invoking an adversary callback} If we have a higher-order interface to
the adversary, we might need to invoke an adversary callback. In this case, not
so much changes with respect to the situation where we invoke regular adversary
code. The adversary can provide a callback as a capability for us to jump to,
either an $\entry$-capability if he wants to protect himself from us or just an
$\exec$ capability if they is not worried about that. However, there is one
scenario that we need to prevent, namely that they constructs the callback
capability to point into the stack, where they may have stored local
capabilities that he should not have access to upon invocation of the callback.
Specifically, he might store return pointers and stack pointers from previous
stack frames there, and illegally use them inside the callback.
